---
title: "Preparing the NKR dataset"
author: "Mary Ann Binuya"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  pdf_document: default
---

# Aim(s)
1. To confirm Netherlands Cancer Registry (NKR) data received meet study’s eligibility criteria
2. To explore NKR data
3. To prepare merged (NKR+CBS) data for analysis. CBS = Centraal Bureau voor de Statistiek, where cause of death are available.
4. To impute missing data
5. To run PREDICT v2.3 on imputed data
6. To check significance of Mammaprint for predicting breast cancer and other/unknown mortality
7. To check proportionality of hazards
8. To check linearity of continuous variables
9. Check interaction/effect modification of Mammaprint

```{r setup, include=FALSE}
#Set libpath
#.libPaths(c("U:/libraries", .libPaths()))

knitr::opts_chunk$set(
  echo = TRUE,
  message = TRUE,
  warning = FALSE,
  fig.retina = 3,
  fig.path = "Output/1 Data Prep/")

#Load libraries
library(haven) #for data imports
library(dplyr) #for data manipulations
library(mice) #for imputations
library(Hmisc) #for labeling
library(survival) #for survival analysis
library(cmprsk) #for competing risk analysis
library(riskRegression) #for fgr
library(lava) #for fgr
library(prodlim) #for fgr

rm(list=ls())
```

# 1. Load raw NKR data
```{r DataloadNKR}
datankr <- read_sav("H:/Analysis/Data/8945_K20298CBKV1.sav")

# Confirm eligibility criteria
  table(datankr$Gesl, useNA = "always") #confirm women only
  table(datankr$Pre_mal, useNA = "always") #confirm no previous malignancies
  table(!is.na(datankr$Incjr)) #confirm diagnosed with invasive BC
  table(datankr$Chir, useNA = "always") #confirm underwent surgery
  all.equal(datankr$Incjr, datankr$Jaarvg) #extra check: Incjr and Jaarvg have exactly the same values

# All study-specific eligibility criteria met.
```

# 2. Explore NKR data
```{r Explore}
# Check distinct records
  n_distinct(datankr$RINPERSOON) #some records come from the same identifier/ID (RINPERSOON) 
  table(datankr$RINPERSOON == "") #few records have missing ID, ignore as cannot match with CBS
  table(datankr$Jaarvg, useNA="always") #there are approx 15,000 patients with BC yearly. we have fewer, ~10,000 per year, because we requested women with early-stage unilateral breast cancer (BC)
  
# Explore records with same ID
  datadup <- datankr %>%
  group_by(RINPERSOON) %>% filter(n() > 1, RINPERSOON != "")
  dim(datadup)[1] #only 2/28 records with repeated RINPERSOON (ID) had Mammaprint data. One repeat was a recurrence, not a first primary breast cancer (BC). We manually exclude this patient in the merging chunk below. The other repeat is handled in further data selection.
  
# Examine other data
  names(select(datankr, contains("dat"))) #there are 3 date variables: "Geboortedatum" "Overlijdensdatum" "Peildatum"
  table(datankr$Vitstat[datankr$Overlijdensdatum != "."], useNA = "always") #295 records with Vitstat=0 (alive) but with death date; use death date not Vitstat for events.

  table(datankr$Oestrrec_stat, as.numeric(datankr$Oestrrec_uitslag), useNA = "always") #ER (estrogen receptor) uses 10% cutoff
  table(datankr$Progrec_stat, as.numeric(datankr$Progrec_uitslag), useNA = "always") #PR (progesterone receptor) uses 10% cutoff
  table(datankr$Mammaprint, datankr$Incjr, useNA = "always") #MammaPrint data was available only from 2011 onwards
  table(datankr$Mammaprint, datankr$Oncotype, useNA = "always") #only four records with both MammaPrint and Oncotype data, not enough to do extra analysis
```

# 3. Select relevant predictors and records (NKR data)
```{r Datasmall}
# Select relevant variables
  datankr_small <- datankr %>%
    select(RINPERSOON, #identifier for CBS matching
      Geboortedatum, #birth date
      Peildatum, #reference/diagnosis date
      Incjr, #year included
      Overlijdensdatum, #death date
      Perfstat, #performance status
      Leeft, #age at diagnosis
      Menopstatus, #menopausal status
      Oestrrec_stat, #ER status
      Progrec_stat, #PR status
      Her2_uitslag, #HER2 status
      ###Ki67 (no variable available, assume unknown)
      Tumorgrootte, #tumor size
      Diffgr, #tumor grade
      Scrndetec, #tumor detected at screening
      Lypos, #number of positive lymph nodes
      starts_with('Chir_type'), #surgery type
      Chemo, #chemotherapy
      starts_with('Chemo_type'), #chemotherapy type
      Horm, #hormone therapy
      starts_with('Horm_type'), #hormone therapy type
      starts_with('Horm_duur'), #hormone therapy length
      Target, #targeted therapy
      starts_with('Target_type'), #targeted therapy type
      Rt, #radiotherapy
      starts_with('Rt_type'), #radiotherapy type
      ###Bisphosphonates (not available, assume based on menopausal status and age)
      Mammaprint,
      starts_with('Cci_'), #Comorbidity data including metastases
      Pt, #pathological tumor size
      Pn, #pathological node involvement
      Pm, #pathological metastases
      Stadiump, #pathological TNM stage
      Meta1, Meta2, Meta3, Meta4, #metastases localization
      Meta_ther, #therapy targeting metastasis
      Meta_ther_type1, Meta_ther_type2, #type of therapy targeting metastasis
      Sec_breast_andermal, #secondary breast tumor
      Sec_andermal, #second primary tumor
      starts_with('Sec_maltumorsoort'), #secondary tumor type
      starts_with('Fu5y_rec'), starts_with('Fu10y_rec')) #events at follow-up including recurrence

#Select relevant records
  datankr_smallmp <- datankr_small %>% filter(Mammaprint %in% c('2', '3')) #with Mammaprint
  dim(datankr_smallmp)[1] #3563 records
  
  datankr_smallER <- datankr_smallmp %>% filter(Oestrrec_stat=='1') #ER-positive, available for all
  dim(datankr_smallER)[1] #3360 records
  
  datankr_smallsize <- datankr_smallER %>% filter(Tumorgrootte!=0) #size >0 (note: size=0 causes the linear predictor/prognostic index of PREDICT's breast cancer specific mortality submodel to be -Inf); retain missing tumor size for imputation
  dim(datankr_smallsize)[1] #3350 records
  
  datankr_smallID <- datankr_smallsize %>% filter(!RINPERSOON %in% c('')) #with ID (needed for CBS matching)
  dim(datankr_smallID)[1] #3324 records

#Recode missing data
  is.na(datankr_smallID) <- datankr_smallID==''
  is.na(datankr_smallID$Tumorgrootte) <- datankr_smallID$Tumorgrootte=='999'
  is.na(datankr_smallID$Progrec_stat) <- datankr_smallID$Progrec_stat=='9'
  is.na(datankr_smallID$Lypos) <- datankr_smallID$Lypos=='98'
  is.na(datankr_smallID$Diffgr) <- datankr_smallID$Diffgr=='9'
  is.na(datankr_smallID$Overlijdensdatum) <- datankr_smallID$Overlijdensdatum=='.'

#Reformat variables
  datankr_formatted <- as.data.frame(unclass(datankr_smallID), stringsAsFactors=TRUE)
  rm(list=setdiff(ls(), c('datankr_formatted', 'datankr')))
  
  datankr_formatted$Geboortedatum = as.Date(datankr_formatted$Geboortedatum, format='%Y%m%d')
  datankr_formatted$Peildatum = as.Date(datankr_formatted$Peildatum, format='%Y%m%d')
  datankr_formatted$Overlijdensdatum = as.Date(datankr_formatted$Overlijdensdatum, format='%Y%m%d')
  
  datankr_formatted$RINPERSOON <- as.character(datankr_formatted$RINPERSOON)
  datankr_formatted$Leeft <- as.numeric(as.character(datankr_formatted$Leeft))
  datankr_formatted$Tumorgrootte <- as.numeric(as.character(datankr_formatted$Tumorgrootte))
  datankr_formatted$Lypos <- as.numeric(as.character(datankr_formatted$Lypos))
  
#Create event and time-to-event variables with cut-off: 2019-12-31

  #Overall survival (OS)
  datankr_formatted$os <- ifelse(datankr_formatted$Overlijdensdatum <= as.Date('2019-12-31'), 1, 0)
  datankr_formatted$os[is.na(datankr_formatted$os)] <- 0
  table(datankr_formatted$os, useNA='always')
  
  #Follow-up time OS (time) in years
  datankr_formatted$time = ifelse(datankr_formatted$os==1,
                                  round(as.numeric(difftime(datankr_formatted$Overlijdensdatum, datankr_formatted$Peildatum))/365.25,2),
                                  round(as.numeric(difftime(as.Date('2019-12-31'), datankr_formatted$Peildatum))/365.25,2))
```

# 3. Load and prepare CBS data
```{r DataloadCBS}
# Link NKR with CBS data until end of 2019 (note: during analysis, CBS codes affected by COVID (2020 onwards) were still being finalized)
  datacbsn011 <- read_sav("H:/Analysis/Data/DO2011V2.sav")
  datacbsn012 <- read_sav("H:/Analysis/Data/130708_DO_2012V1.sav")
    datacbsn011 <- datacbsn011 %>% select("rinpersoon", "ovljr", "primoorz")
    datacbsn012 <- datacbsn012 %>% select("rinpersoon", "ovljr", "primoorz")
    datacbs1 <- rbind(datacbsn011, datacbsn012)
    datacbs1 <- datacbs1 %>% rename(RINPERSOON = rinpersoon, STATJAAR = ovljr, UCCODE = primoorz)
    rm(list = ls(pattern = "datacbsn0"))
    
  datacbs2013 <- read_sav("H:/Analysis/Data/DOODOORZ2013TABV2.sav") #database more simplified from 2013 onwards
  datacbs2014 <- read_sav("H:/Analysis/Data/DOODOORZ2014TABV1.sav")
  datacbs2015 <- read_sav("H:/Analysis/Data/DOODOORZ2015TABV2.sav")
  datacbs2016 <- read_sav("H:/Analysis/Data/DOODOORZ2016TABV2.sav")
  datacbs2017 <- read_sav("H:/Analysis/Data/DOODOORZ2017TABV2.sav")
  datacbs2018 <- read_sav("H:/Analysis/Data/DOODOORZ2018TABV2.sav")
  datacbs2019 <- read_sav("H:/Analysis/Data/DOODOORZ2019TABV2.sav")
    datacbs2 <- do.call("rbind", mget(ls(pattern = "ˆdatacbs20")))
    rm(list = ls(pattern = "datacbs20"))
    
# Check redundant records
  n_occur <- data.frame(table(datacbs2$RINPERSOON))
  n_occur[n_occur$Freq > 1, ] #there are two redundant records
    rm(n_occur)

  ## Explore the two redundant records:
    table(datacbs2$STATJAAR[datacbs2$RINPERSSON == "[deleted_ID1]"], datacbs2$UCCODE[datacbs2$RINPERSSON == "[deleted_ID1]"])
      #recoded from R99 (ill-defined/unknown cause of mortality in 2015) to
      #F03 (unspecified dementia in 2017)
    table(datacbs2$STATJAAR[datacbs2$RINPERSSON == "[deleted_ID2]"], datacbs2$UCCODE[datacbs2$RINPERSSON == "[deleted_ID2]"])
      #recoded from R99 (ill-defined/unknown cause of mortality in 2016) to
      #C349 (malignant neoplasm of unspecified part of bronchus or lung in 2018)
    
# Select to reflect the most up-to-date death codes
  datacbs2.uniq <- subset(datacbs2, !(RINPERSOON == "[deleted_ID1]" & UCCODE == "R99"))
  datacbs2.uniq <- subset(datacbs2.uniq, !(RINPERSOON == "[deleted_ID2]" & UCCODE == "R99"))
  
  datacbs2.uniq2 <- data.frame(datacbs2.uniq) %>% select("RINPERSOON", "STATJAAR", "UCCODE")
  
  datacbs <- rbind(datacbs1, datacbs2.uniq2)
```

# 3. Combine NKR and CBS data, and explore data
```{r Discrepancies}
# Combine NKR and CBS data
  datamerge <- left_join(datankr_formatted, datacbs, by = c(RINPERSOON = "RINPERSOON"))

# Explore discrepancies between NKR and CBS data

  ## 1 NKR death not in CBS (retain with unknown as cause of death)
  datamerge$RINPERSOON[datamerge$os == 1 & is.na(datamerge$UCCODE)]

    ### Check that it is not in original CBS databases
    table(datacbs1$UCCODE[datacbs1$RINPERSOON == "[deleted_ID3]"], useNA = "always") #confirmed
    table(datacbs2$UCCODE[datacbs2$RINPERSOON == "[deleted_ID3]"], useNA = "always") #confirmed

  datamerge$UCCODE[datamerge$RINPERSOON == "[deleted_ID3]"] <- "R99" #retain record, with cause of death=unknown 

  ## 2 CBS deaths not in NKR
  datamerge$RINPERSOON[datamerge$os == 0 & !is.na(datamerge$UCCODE)]
  
    ### One record [deleted_ID4], very likely recurrence (see above notes on duplicate records), exclude
    datamerge <- subset(datamerge, RINPERSOON != "[deleted_ID4]") #
  
    ### One record [deleted_ID5]: retain but add approximate death date (year) as only 2017 is indicated in CBS.
    datamerge$os[datamerge$RINPERSOON == "[deleted_ID5]"] <- 1
    
    datamerge$Overlijdensdatum[datamerge$RINPERSOON == "[deleted_ID5]"] <- as.Date("2017-12-31") #approx death date
    datamerge$time[datamerge$RINPERSOON == "[deleted_ID5]"] <- round(as.numeric(difftime(
      datamerge$Overlijdensdatum[datamerge$RINPERSOON == "[deleted_ID5]"],
      datamerge$Peildatum[datamerge$RINPERSOON == "[deleted_ID5]"]))/365.25, 2)
    
  rm(datacbs1, datacbs2, datacbs2.uniq, datacbs2.uniq2)

# Unique causes of death
  n_occur <- data.frame(table(datamerge$UCCODE[datamerge$os == 1]))
  sum(n_occur$Freq) #145 deaths
  n_occur[order(-n_occur$Freq), ] #49 unique ICD10 codes
  rm(n_occur)
  
# Create event: competing risk
  datamerge$oscr <- NA
  datamerge$oscr <- ifelse(datamerge$os == 1 & datamerge$UCCODE == "C509", 1, ifelse(datamerge$os == 0, 0, 2)) #C509 = BC-related death
  table(datamerge$oscr, useNA = "always")  #65 BC-related deaths overall
  round(prop.table(table(datamerge$oscr, useNA = "always")) * 100, 1) 
  table(datamerge$Mammaprint, datamerge$oscr, useNA = "always")
```

#3. Prepare combined data for imputation
```{r Impprep}
# Check format/reformat variables for imputation

  ## grisk (MammaPrint/70-gene signature)
  datamerge$grisk <- ifelse(datamerge$Mammaprint == 2, 0, 1)
  datamerge$grisk <- as.factor(datamerge$grisk)
  table(datamerge$grisk, useNA = "always")

  ## grade
  table(datamerge$Diffgr, useNA = "always") #grade not allowed to be unknown in PREDICT online tool, impute in a standard manner

  ## her2
  datamerge$her2 <- NA #her2 allowed to be unknown in PREDICT online tool and source code, but impute here nonetheless
  datamerge$her2 <- ifelse(datamerge$Her2_uitslag == "3", 1,
                           ifelse(datamerge$Her2_uitslag == "0" | datamerge$Her2_uitslag == "1", 0, datamerge$her2))
  datamerge$her2 <- as.factor(datamerge$her2)
  table(datamerge$her2, useNA = "always")

  ## chemo generation
  datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")] <-
    lapply(datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")], as.character)
  datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")] <-
    lapply(datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")], function(x)
      replace(x, grepl("426000|451000|452000|454000", x), 2))
  datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")] <-
    lapply(datamerge[, c("Chemo_type1", "Chemo_type2", "Chemo_type3")], function(x)
      replace(x, grepl("420000|421000|427000|466000|L01CD01|L01CD02", x), 3))
  datamerge$generation <- pmax(datamerge$Chemo_type1, datamerge$Chemo_type2, datamerge$Chemo_type3, na.rm = TRUE)
  datamerge$generation <- ifelse(is.na(datamerge$generation), 0, datamerge$generation)
  datamerge$generation <- as.factor(datamerge$generation)
  table(datamerge$Chemo, datamerge$generation, useNA = "always") #sanity check that chemo codes are appropriate
  
  ## horm
  table(datamerge$Horm, useNA = "always") 

  ## traz (trastuzumab/anti-her2 therapy)
  datamerge$traz <- ifelse(datamerge$Target_type1 == "L01XC03" |
                             datamerge$Target_type2 == "L01XC03" |
                             datamerge$Target_type3 == "L01XC03", 1, 0)
  datamerge$traz <- ifelse(is.na(datamerge$traz), 0, datamerge$traz)
  table(datamerge$traz, useNA = "always")

# Create other variables necessary for PREDICT risk calculation
  
  ## bisphophonates - derive based on menopausal status and age
  datamerge$bis <- ifelse(datamerge$Menopstatus == "2" | datamerge$Menopstatus == "4", 1,
                          ifelse(datamerge$Menopstatus == "1" | datamerge$Menopstatus == "3", 0,
                                 ifelse(datamerge$Menopstatus == "9" & datamerge$Leeft > 50, "1", 0)))
  datamerge$bis <- as.factor(datamerge$bis)
  table(datamerge$bis, useNA = "always")
  
  ## ki67 - assume unknown for all
  datamerge$ki67 <- 9
  
# Prepare final data
  dataimp <- datamerge %>%
    select(Leeft,
           Scrndetec,
           Tumorgrootte, Pt,
           Diffgr,
           Lypos, Pn,
           Oestrrec_stat,
           Progrec_stat,
           her2,
           ki67,
           generation,
           Horm,
           traz,
           bis,
           grisk,
           time,
           os,
           oscr) %>%
  rename(age.start = Leeft,
         screen = Scrndetec,
         size = Tumorgrootte,
         grade = Diffgr,
         nodes = Lypos,
         er = Oestrrec_stat,
         pr = Progrec_stat,
         horm = Horm)
```

# 4. Impute data - MICE imputation
```{r Imp}
# Check pattern of missingness
  md.pattern(dataimp) #assume MAR; fraction of missing information: 144/3323 = ~5%, 10 imputations should be sufficient

# Compute baseline cumhaz
  dataimp$eventbc <- ifelse(dataimp$oscr == 1, 1, 0)
  dataimp$eventoth <- ifelse(dataimp$oscr == 2, 1, 0)
  
  dataimp$cumhaz_bc <- nelsonaalen(dataimp, time, eventbc) # define marginal baseline cumhaz
  dataimp$cumhaz_oth <- nelsonaalen(dataimp, time, eventoth) # define marginal baseline cumhaz

# Imputation
  ini <- mice(dataimp, maxit = 0, seed = 202309)
  ini$nmis #number missing per variable: pr=2, her2=12, nodes=15, size=34, grade=81
  
  vis <- c("pr", "her2", "nodes", "size", "grade") #define order of imputation from lowest to highest number of missing
  
  method <- ini$method #check appropriate methods are used in imputation
  method #pr=logreg, her2=logreg, nodes=pmm, size=pmm, grade=polyreg
  
  ## Retain as predictors: survival time, relevant status indicator, baseline cumhaz, and relevant covariates
  pred_bc <- ini$predictorMatrix
  pred_bc[, "bis"] <- 0 #derived variable
  pred_bc[, "os"] <- 0
  pred_bc[, "oscr"] <- 0 #retain eventbc
  pred_bc[, "eventoth"] <- 0
  pred_bc[, "cumhaz_oth"] <- 0 #retain cumhaz_bc
  pred_bc #keep as predictors: 
  
  pred_oth <- ini$predictorMatrix
  pred_oth[, "bis"] <- 0 #exclude, derived variable
  pred_oth[, "os"] <- 0
  pred_oth[, "oscr"] <- 0 #retain eventoth
  pred_oth[, "eventbc"] <- 0
  pred_oth[, "cumhaz_bc"] <- 0 #retain cumhaz_oth
  pred_oth
  
  ## Constrain imputation of size and nodes ased on pathological size and nodes classification
  post <- ini$post
  post["size"] <- "
  imp[[j]][data$Pt[!r[,j]] == '1A', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == '1A', i], c(0, 5))
  imp[[j]][data$Pt[!r[,j]] == '1B', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == '1B', i], c(5, 10))
  imp[[j]][data$Pt[!r[,j]] == '1C', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == '1C', i], c(10, 20))
  imp[[j]][data$Pt[!r[,j]] == '2', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == '2', i], c(20, 50))
  imp[[j]][data$Pt[!r[,j]] == '3', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == '3', i], c(50, 100))
  imp[[j]][data$Pt[!r[,j]] == 'X', i] <- squeeze(imp[[j]][data$Pt[!r[,j]] == 'X', i], c(0, 100))"
  post["nodes"] <- "
  imp[[j]][data$Pn[!r[,j]] == '0S', i] <- squeeze(imp[[j]][data$Pn[!r[,j]] == '0S', i], c(0, 0))
  imp[[j]][data$Pn[!r[,j]] == '2A', i] <- squeeze(imp[[j]][data$Pn[!r[,j]] == '2A', i], c(4, 9))
  imp[[j]][data$Pn[!r[,j]] == 'X', i] <- squeeze(imp[[j]][data$Pn[!r[,j]] == 'X', i], c(0, 25))"
  
  ## Impute
  imp_bc <- mice(dataimp, m = 10, method = method, visitSequence = vis, predictorMatrix = pred_bc, post = post, seed = 202309)
  sapply(complete(imp_bc), function(x) sum(is.na(x))) #check no more missing
  plot(imp_bc) #check convergence; convergence is better with m=10 than m=5; retain m=10.
  densityplot(imp_bc, ~pr+her2 + nodes + size + grade) #check distribution of imputed values
  
  imp_oth <- mice(dataimp, m = 10, method = method, visitSequence = vis, predictorMatrix = pred_oth, post = post, seed = 202309)
  sapply(complete(imp_oth), function(x) sum(is.na(x))) #check no more missing
  plot(imp_oth) #check convergence is ok
  densityplot(imp_oth, ~pr+her2 + nodes + size + grade)
  
rm(ini, pred_bc, pred_oth, post, method, vis)
```

#5. Run PREDICT on imputed dataset(s)
```{r PREDICT}
# Load PREDICT version 2.3 function
  source("Functions/predictv23_batchrunner.R")

# Run PREDICT on imp_bc
  datapredimp_bc <- complete(imp_bc, action = "long", include = TRUE) # complete imputed data
  
  datapredimp_bc[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")] <-
    as.numeric(as.matrix(datapredimp_bc[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")])) #format variables to accommodate PREDICT function; does not change any individual values
  
  datapredimp_bc <- ukpredict(datapredimp_bc, year = 5, age.start = age.start, screen = screen,
                              size = size, grade = grade, nodes = nodes, er = er, pr = pr, her2 = her2, ki67 = ki67,
                              generation = generation, horm = horm, traz = traz, bis = bis)
  
  datapredimp_bc[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")] <-
    lapply(datapredimp_bc[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")], FUN = as.factor) #format variables back
  
  imppred_bc <- as.mids(datapredimp_bc)

# Run PREDICT on imp_oth
  datapredimp_oth <- complete(imp_oth, action = "long", include = TRUE) #complete imputed data
  
  datapredimp_oth[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")] <-
    as.numeric(as.matrix(datapredimp_oth[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")])) #format variables to accommodate PREDICT function; does not change any individual values
  
  datapredimp_oth <- ukpredict(datapredimp_oth, year = 5, age.start = age.start, screen = screen,
                               size = size, grade = grade, nodes = nodes, er = er, pr = pr, her2 = her2, ki67 = ki67,
                               generation = generation, horm = horm, traz = traz, bis = bis)
  
  datapredimp_oth[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")] <-
    lapply(datapredimp_oth[, c("er", "pr", "her2", "ki67", "generation", "horm", "traz", "bis")], FUN = as.factor) #format variables back
  
  imppred_oth <- as.mids(datapredimp_oth)
  rm(list = setdiff(ls(), c("datapredimp_bc", "imppred_bc", "datapredimp_oth", "imppred_oth")))

# save.image('WS_1_Data_Prep_NKR.RData')
```

# 6. To check significance of Mammaprint for predicting breast cancer and other/unknown mortality
```{r Outcomes}
# PREDICT considers two competing events in its calculation of overall survival: breast cancer and other mortality. We confirmed in MINDACT dataset that Mammaprint is only a significant predictor for breast cancer mortality. We confirm this again in the NKR dataset.

# Cause-specific models:
  ## Breast cancer
  fit_bcss <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk))
  summary(pool(fit_bcss)) #MammaPrint (genomic risk) a significant predictor for breast cancer mortality, consistent with MINDACT results.
  
  fit_bcss_m <- imppred_bc %>%
  with(coxph(Surv(time, eventbc) ~ offset(pirx) + grisk))
  summary(pool(fit_bcss_m)) #significance retained adjusting for clinical risk (via PREDICT)
    #grisk = 0.67, same effect as in MINDACT dataset
  
  fit_bcss_m <- imppred_bc %>%
  with(coxph(Surv(time, eventbc) ~ pirx + grisk))
  summary(pool(fit_bcss_m)) #significance retained adjusting for clinical risk (via PREDICT)

  ## Other
  fit_oth <- imppred_oth %>% with(coxph(Surv(time, eventoth) ~ grisk))
  summary(pool(fit_oth)) #MammaPrint (genomic risk) NOT a significant predictor for other/unknown mortality, consistent with MINDACT results.
  
  fit_oth_m <- imppred_oth %>% with(coxph(Surv(time, eventoth) ~ mi + grisk))
  summary(pool(fit_oth_m)) #still NOT a significant predictor after adjustment for PREDICT other-mortality specific prognostic index

  rm(list = ls(pattern = "ˆfit"))

# Sanity check using subdistribution hazard models
  ## Breast cancer
  fgr_res <- fgr_res_m <- list()

  for (i in 1:10) {
    df <- datapredimp_bc %>% filter(.imp == i)
    fgr_bcss <- FGR(Hist(time, oscr) ~ grisk, cause = 1, data = df)
    fgr_res[[i]] <- summary(fgr_bcss)$coef
    fgr_bcss_m <- FGR(Hist(time, oscr) ~ pirx + grisk, cause = 1, data = df)
    fgr_res_m[[i]] <- summary(fgr_bcss_m)$coef
    }

  fgr_res #similar result as cause-specific model
  fgr_res_m #similar result as cause-specific model
  
  rm(list = ls(pattern = "ˆfgr"))
  
  ## Other
  fgr_res <- fgr_res_m <- list()
  
  for (i in 1:10) {
    df <- datapredimp_bc %>%
      filter(.imp == i)
      fgr_oths <- FGR(Hist(time, oscr) ~ grisk, cause = 2, data = df)
      fgr_res[[i]] <- summary(fgr_oths)$coef
      fgr_oths_m <- FGR(Hist(time, oscr) ~ mi + grisk, cause = 2, data = df)
      fgr_res_m[[i]] <- summary(fgr_oths_m)$coef
      }

  fgr_res #similar result as cause-specific model
  fgr_res_m #similar result as cause-specific model

  rm(list = ls(pattern = "^fgr"))
  rm(df)
  
# Limit analysis to prediction of prediction of breast cancer mortality.
```

# 7. Check proportionality of hazards
```{r PH}
zph <- zph_full <- list()

for (i in 1:10) {
  dat <- datapredimp_bc %>%
    filter(.imp == i) %>%
    select(time, eventbc,
           grisk, pirx, age.start, size, grade, nodes, pr, her2, generation, horm, traz, bis)

  fit_bc <- coxph(Surv(time, eventbc) ~ grisk, data = dat)
  res <- cox.zph(fit_bc, transform = "identity")
  zph[[i]] <- res
  
  fit_bc_full <- coxph(Surv(time, eventbc) ~ ., data = dat) #include all covariates
  res_full <- cox.zph(fit_bc_full, transform = "identity")
  zph_full[[i]] <- res_full
}

rand <- sample(1:10, 1) #show result only of the ith random imputed data

zph[[rand]] #grisk hazard is proportional over time grisk P=0.22, global P=0.22 
zph_full[[rand]] #grisk hazard is proportional over time grisk P=0.18, pirx=0.50, global P=0.42 
plot(res_full, resid = FALSE)

  rm(list = ls(pattern = "^fit"))
  rm(list = ls(pattern = "^zph"))
  rm(list = ls(pattern = "^res"))
  rm(i, rand)

# Based on the Schoenfeld test, the PH assumption holds for MammaPrint/genomic risk (grisk), PREDICT's prognostic index (pirx), and other covariates relevant for prediction of risk for breast cancer mortality.
```

# 8. Check linearity of continuous variables
```{r Lin}
# To assess linearity, we fit a linear and non-linear term and check their significance. We do this using pspline() in coxph.

fit_pirx <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ pspline(pirx))) #if df=0 is specified, pspline will choose the best df based on AIC
summary(fit_pirx) #linear

  rm(list = ls(pattern = "^fit"))
```

# 9. Check interaction/effect modification of Mammaprint
```{r Effectmodification}
fit_gpirx <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * pirx))
summary(pool(fit_gpirx))  #no significant interaction

fit_gage <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * age.start))
summary(pool(fit_gage))  #some interaction but not statistically significant at alpha=5%

fit_gscreen <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * screen))
summary(pool(fit_gscreen))  #no significant interaction

fit_gsize <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * size))
summary(pool(fit_gsize))  #no significant interaction

fit_ggrade <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * grade))
summary(pool(fit_ggrade))  #no significant interaction

fit_gnodes <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * nodes))
summary(pool(fit_gnodes))  #no significant interaction

fit_gpr <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * pr))
summary(pool(fit_gpr))  #no significant interaction

fit_gher2 <- imppred_bc %>% with(coxph(Surv(time, eventbc) ~ grisk * her2))
summary(pool(fit_gher2))  #no significant interaction, but too few patients in HER2+ due to selection

  rm(list = ls(pattern = "^fit"))
```

